---
title: "MTG price archive — first exploration"
author: "Michael"
date: today
format:
  html:
    toc: true
    code-fold: show
---

# Setup

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(arrow)       # open_dataset over S3 Parquet
  library(dplyr)       # tidyverse verbs on Arrow tables
  library(ggplot2)
  library(scales)
  library(lubridate)
  library(here)
})

source(here::here("R", "theme.R"))

bucket <- Sys.getenv("MTG_S3_BUCKET", "mtg-scrape-unwindgames")
region <- Sys.getenv("AWS_REGION",    "us-east-1")
```

# Open the Parquet dataset

Reads directly from S3 without pulling the data locally. `arrow` pushes
filter/select down to the file reader, so the Parquet columnar format pays
off on every query.

```{r}
prices <- open_dataset(
  sources = sprintf("s3://%s/prices/?region=%s", bucket, region),
  format  = "parquet"
)

prices
```

# Row counts and shape

Total row count across the 90-day archive:

```{r}
prices |>
  summarise(rows = n()) |>
  collect()
```

Distribution across vendor × kind × finish:

```{r}
prices |>
  count(vendor, kind, finish, sort = TRUE) |>
  collect()
```

# Daily observation volume

Does every date partition have roughly the same row count, or are some
providers dropping out on older dates?

```{r daily-volume, fig.cap = "Daily observation volume by vendor, 90-day window"}
daily <- prices |>
  filter(kind == "retail") |>
  count(date, vendor) |>
  collect()

daily |>
  ggplot(aes(date, n, color = vendor)) +
  geom_line(linewidth = 0.4) +
  scale_y_continuous(labels = label_comma()) +
  labs(
    title    = "Daily price observations per vendor",
    subtitle = "Retail only; shows how completely each vendor reports across the 90-day window",
    x = NULL, y = NULL,
    color = NULL,
    caption = "Source: MTGJSON AllPrices archive"
  ) +
  theme_mtg()
```

# Top-20 TCGplayer trajectories

For a quick sanity check, plot price trajectories for the 20 most expensive
non-foil cards on TCGplayer as of the latest archive date.

```{r tcgplayer-top20, fig.cap = "Top 20 TCGplayer non-foil cards by latest market price, log scale"}
#| cache: true

snapshot <- prices |>
  filter(vendor == "tcgplayer", kind == "retail", finish == "normal") |>
  select(card_uuid, date, price) |>
  collect()

latest_day <- max(snapshot$date)

top20 <- snapshot |>
  filter(date == latest_day) |>
  slice_max(price, n = 20) |>
  pull(card_uuid)

snapshot |>
  filter(card_uuid %in% top20) |>
  ggplot(aes(date, price, group = card_uuid)) +
  geom_line(alpha = 0.45, linewidth = 0.4, color = mtg_palette["black"]) +
  scale_y_log10(labels = label_dollar()) +
  labs(
    title    = sprintf("TCGplayer retail — top 20 as of %s", latest_day),
    subtitle = "Non-foil only; one faint line per card; log y-axis",
    x = NULL, y = NULL,
    caption = "Source: MTGJSON AllPrices archive"
  ) +
  theme_mtg()
```

# Next steps

- Join the card dimension (`dimension/mtgjson/allprintings/`) to attach
  set, rarity, color identity, and Reserved List flag.
- Add EDHREC / tournament play-rate signal once the ingest is wired.
- Build a reprint-event feature table and model price shocks around
  announcement dates.
